---
title: "AHS-2 lupus study 2"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "tableone", "emmeans", "kableExtra", "Gmisc")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, pattern, ...) {
  loc <- grep(pattern, names(df), ...)
  if (length(loc) == 0) warning("There are no such variables")
  else return(data.frame(loc = loc, varname = names(df)[loc]))
}

# Read data: n = 93,467 and 111 variables
filepath <- "./data/lupus-initial-dataset-v1-2022-04-25.csv"
lupus0  <- read_csv(filepath)

# Additional n-6 variables
filepath <- "./data/lupus-initial-dataset-v2-2022-06-03.csv"
lupus00  <- read_csv(filepath)

# Cod liver oil
filepath <- "./data/lupus-dataset-hhf3-codliver-v3-2022-06-29.csv"
lupus00a  <- read_csv(filepath)

# HHF3
filepath <- "./data/lupus-hhf3-return-date.csv"
lupus00b <- read_csv(filepath)

# Merge all files
lupus0 <- lupus0 %>% 
  inner_join(select(lupus00, analysisid, starts_with("p182"), starts_with("p204"))) %>% 
  left_join(select(lupus00a, analysisid, codliver, hhf3)) %>% 
  left_join(select(lupus00b, analysisid, hhf3_returned))

# Include non-Hispanic White or Black
# Exclude age < 30
# Exclude missing gender, diet pattern, education, smoking and BMI
# Exclude extreme kcal <500 or >4500
lupus <- lupus0 %>%
  mutate(
    age = ifelse(age < 30, NA, age),
    agecat = cut(age, breaks = c(0, 40, 60, Inf), right = FALSE),
    agecat = factor(agecat, labels = c("30-39", "40-59", ">=60")),
    sex = factor(sex, labels = c("Female", "Male")),
    black = case_when(
      ethyou == "01" ~ 0,
      ethyou %in% c("02", "03", "04", "05", "39", "40", "63", "74", "77", "96") ~ 1),
    black = factor(black, labels = c("White", "Black")),
    smkcat  = case_when(
      smokenow == 1 ~ 3,
      smoke > 1 & is.na(smokenow) ~ 2,
      smoke == 1 ~ 1
    ),
    smkcat  = factor(smkcat, labels = c("Never", "Past", "Current")),
    smkever = factor(smkcat, labels = c("Never", "Ever", "Ever")),
    educat3 = cut(educyou, breaks = c(0, 3, 6, 9)),
    educat3 = factor(educat3, labels = c("HS or less", "Some college", "Col grad")),
    # take_vd = ifelse(!is.na(vitd) & vitd == 2, 1, 0),
    # take_vd = factor(take_vd, labels = c("No", "Yes")),
    # take_fo = ifelse(!is.na(fishoil) & fishoil == 2, 1, 0),
    take_fo = ifelse((!is.na(fishoil) & fishoil == 2) | (!is.na(codliver) & codliver == 1), 1, 0),
    take_fo = factor(take_fo, labels = c("No", "Yes")),
    bmicat  = cut(bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
    bmicat  = factor(bmicat, labels = c("Normal", "Overweight", "Obese")),
    vegstat = factor(vege_group_gen_bl, levels = c("vegan", "lacto", "pesco", "semi", "nonveg")),
    vegstat3 = vegstat,
    prev_sle = ifelse(!is.na(sle) & sle == 2, 1, 0),
    prev_sle = factor(prev_sle, labels = c("No", "Yes"))) %>% 
  drop_na(sex, age, black, vege_group_gen_bl, smkever, educat3, bmi) %>% 
  filter(kcal >= 500 & kcal <= 4500)

levels(lupus$vegstat) <- c("Vegans", "Lacto-ovo", "Pesco", "Semi", "Non-veg")
levels(lupus$vegstat3) <- c("Vegetarians", "Vegetarians", "Pesco", "Non-veg", "Non-veg")
```

## Dataset

* Filename: `lupus-initial-dataset-v1-2022-04-25.csv`
  * Data received on 4/25/2022
  * Includes *n* = `r nrow(lupus0)` subjects and `r ncol(lupus0)` variables
  * No imputations
  
* To create an analytic file:
  * Included only non-Hispanic White or Black
  * Excluded if age at baseline < 30 years
  * Excluded extreme energy intake of <500 or >4500 kcal
  * Excluded any subjects with missing gender, education, smoking history, dietary pattern and BMI
  * This resulted in *n* = `r nrow(lupus)` subjects (`r nrow(lupus0) - nrow(lupus)` subjects were excluded)

* **For now, both males and females are included in the analytic dataset**

## Outcome

* Prevalent cases of SLE were identified using the baseline questionnaire
  * Includes only those who "have been treated for SLE in the last 12 months" at baseline
  * ("Years since the 1st diagnosis" was not used, following the case definition in the manuscript)
* There were `r sum(lupus$prev_sle == "Yes")` prevalent cases (`r round(sum(lupus$prev_sle == "Yes") / nrow(lupus) * 100, 2)`%) of SLE
  
## Descriptive table

* Descriptive table stratified by cases/non-cases
  * Variables were categorized following the original manuscript
  * Use of fish oil supplement (``take_fo``) now includes cod liver oil

```{r echo = FALSE}
table_vars <- c("age", "agecat", "black", "sex", "smkever", "educat3", "vegstat3", "take_fo", "bmi", "bmicat", "kcal")
lupus %>% CreateTableOne(table_vars, strata = "prev_sle", data = .) %>%
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()
```

## Diagnosis of SLE and use of fish oil supplement

* Among those who were diagnosed with SLE and takes fish oil supplement, a crosstab was created between years since diagnosis and the duration of fish oil supplement use:

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(gtsummary)
sle_lab  <- paste(c("<5", "5-9", "10-14", "15-19", "20+"), "years ago")
supp_lab <- c("0-1 year", "2-4 years", "5-9 years", "10+ years")
lupus %>%
  filter(prev_sle == "Yes", sley > 0, fishoily > 0) %>% 
  transmute(sley = factor(sley, labels = sle_lab),
            fishoily = factor(fishoily, labels = supp_lab)) %>% 
  tbl_cross(
    row = sley,
    col = fishoily,
    percent = "none",
    label = list(sley ~ "Diagnosed with SLE", fishoily ~ "Fish oil supplement use: For how long")
  )  %>% 
  as_kable()
```

## Energy-adjusted fatty acid intake

* The nutrient intake data include the following fatty acids
  * Each FA has both dietary and supplement intake

```{r echo = FALSE}
# Nutrients
fa <- tribble(
  ~fa_name, ~lipid, ~note,
  "ALA", "p183", "Omega-3",
  "SDA", "p184", "Omega-3",
  "EPA", "p205", "Omega-3",
  "DPA", "p225", "Omega-3",
  "DHA", "p226", "Omega-3",
  "LA",  "p182", "Omega-6",
  "AA",  "p204", "Omega-6"
)

fa %>% knitr::kable()
```

```{r echo = FALSE}
# Energy-adjust with zero partition
# By default, variables are log-transformed (excluding zeros)
kcal_adjust <- function(var, kcal, log=TRUE){
  if (missing(var))
    stop("Need to specify variable for energy-adjustment.")
  if (missing(kcal))
    stop("Need to specify kcal intake.")
  df <- data.frame(y = var, ea_y = var, kcal = kcal)
  count_negative <- sum(df$y < 0, na.rm=TRUE)
  if (count_negative > 0)
    warning("There are negative values in variable.")
  if(log) df$y[df$y > 0 & !is.na(df$y)] <- log(df$y[df$y > 0 & !is.na(df$y)])
  mod <- lm(y ~ kcal, data=df[df$y != 0, ])
  if(log){
    ea <- exp(resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE))
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  else{
    ea <- resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE)
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  return(df$ea_y)
}

# Create new variables
omega_vars         <- fa$lipid
omega_diet_vars    <- paste0(omega_vars, "diet")
omega_supp_vars    <- paste0(omega_vars, "supp")

omega_vars_ea      <- paste0(omega_vars, "_ea")
omega_diet_vars_ea <- paste0(omega_diet_vars, "_ea")

# Energy-adjust only dietary component (not supplement)
lupus[omega_diet_vars_ea] <- lapply(lupus[omega_diet_vars], kcal_adjust, kcal = lupus$kcal)

# Add energy-adjusted dietary intake and supplement
calc_total_ea <- function(var) rowSums(lupus[c(paste0(var, "diet_ea"), paste0(var, "supp"))])
lupus[omega_vars_ea] <- lapply(omega_vars, calc_total_ea)

# Sum omega fa energy-adjusted values
lupus$n3pfa_ea    <- rowSums(lupus[c("p183_ea", "p184_ea", "p205_ea", "p225_ea", "p226_ea")])
lupus$n6pfa_ea    <- rowSums(lupus[c("p182_ea", "p204_ea")])
lupus$p205p226_ea <- rowSums(lupus[c("p205_ea", "p226_ea")])

# Sum omega fa energy-adjusted dietary values
lupus$n3pfa_diet_ea    <- rowSums(lupus[c("p183diet_ea", "p184diet_ea", "p205diet_ea", "p225diet_ea", "p226diet_ea")])
lupus$n6pfa_diet_ea    <- rowSums(lupus[c("p182diet_ea", "p204diet_ea")])
lupus$p205p226_diet_ea <- rowSums(lupus[c("p205diet_ea", "p226diet_ea")])

# Calculate ratio of FA
lupus <- lupus %>% 
  mutate(o3_o6 = n3pfa_ea / n6pfa_ea,
         p183_o6 = p183_ea / n6pfa_ea,
         p205p226_o6 = p205p226_ea / n6pfa_ea)
```

* For each FA, its dietary intake was energy-adjusted by the residual method, while partitioning zero intake ([Jaceldo-Siegl et al., 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3433053/)). Due to its highly right-skewed distribution, non-zero intake was log-transformed before being regressed on the total energy intake. Residuals were added by the mean of log and then back-transformed to obtain energy-adjusted dietary intake. Energy-adjusted dietary intake was added with (unadjusted) supplement intake to form energy-adjuste total intake.
  * Energy-adjusted total omega-3 fatty acid intake was calculated as the sum of the following energy-adjusted FAs: ALA, SDA, EPA, DPA and DHA.
  * Energy-adjusted total omega-6 fatty acid intake was calculated as the sum of the following energy-adjusted FAs: LA, AA
  
* Distributions of energy-adjusted ALA, EPA + DHA, total omega-3, and total omega-6 are highly right-skewed:

```{r echo = FALSE, fig.width = 10, fig.height = 3}
# Omega-3 and -6 fatty acids
table_vars <- c("p183_ea", "p205p226_ea", "n3pfa_ea", "n6pfa_ea")

# Check distribution
lupus %>% 
  select(all_of(table_vars)) %>% 
  pivot_longer(p183_ea:n6pfa_ea, names_to = "var", values_to = "value") %>% 
  mutate(var = factor(var, levels = table_vars))  %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~var, scales = "free", ncol = 4) +
  labs(x = "Energy-adjusted intake (gram/day)")
```

* Density plots by prevalent SLE status are shown below. Note that the x-axis is in a pseudo-log scale:

```{r echo = FALSE, fig.width = 10, fig.height = 3.5}
# Density plot by SLE status
lupus %>% 
  select(all_of(table_vars), prev_sle) %>% 
  pivot_longer(p183_ea:n6pfa_ea, names_to = "var", values_to = "value") %>% 
  mutate(var = factor(var, levels = table_vars))  %>% 
  ggplot(aes(x = value, fill = prev_sle)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(trans="pseudo_log") +
  facet_wrap(~var, scales = "free", ncol = 4) +
  labs(x = "Energy-adjusted intake (gram/day)", fill = "Prevalent SLE") +
  theme(legend.position = "bottom")
```

* A Spearman correlation matrix among energy-adjusted ALA, EPA + DHA, total omega-3, and total omega-6 is shown below
  * ALA is moderately positively correlated with omega-3 and omega-6 fatty acids.
  
```{r echo = FALSE}
# Check Spearman correlations
lupus %>% 
  select(all_of(table_vars)) %>% 
  cor(method = "spearman") %>% 
  round(3) %>% 
  knitr::kable()
```

* Median (IQR) intake (gram/day) of energy-adjusted ALA, EPA + DHA, total omega-3, and total omega-6 by cases/non-cases:
  * For EPA + DHA and total omega-3, those with SLE had significantly higher intake than those without SLE.

```{r echo = FALSE}
# Table by SLE status
table_vars <- c("p183_ea", "p205p226_ea", "p205p226_diet_ea", "n3pfa_ea", "n3pfa_diet_ea", "n6pfa_ea")

lupus %>% 
  set_column_labels(prev_sle          = "Prevalent SLE", 
                    p183_ea           = "ALA",
                    p205p226_ea       = "DHA + EPA",
                    p205p226_diet_ea  = "DHA + EPA dietary",
                    n3pfa_ea          = "Omega-3",
                    n3pfa_diet_ea     = "Omega-3 dietary",
                    n6pfa_ea          = "Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = prev_sle, 
                        header_count = "(n = %s)",
                        continuous_fn = describeMedian,
                        digits = 2,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) energy-adjusted intake of fatty acids (gram/day)",
            tfoot = "P-values were from Mann-Whitney tests")
```

* Proportions of (energy-adjusted) dietary intake out of the total intake (dietary + supplement) are shown below. 
  * Over 90% of the total intake were from dietary intake, not supplements.

```{r echo = FALSE}
# Sum omaga fa energy-adjusted dietary values
lupus$n3pfa_diet_ea    <- rowSums(lupus[c("p183diet_ea", "p184diet_ea", "p205diet_ea", "p225diet_ea", "p226diet_ea")])
lupus$n6pfa_diet_ea    <- rowSums(lupus[c("p182diet_ea", "p204diet_ea")])
lupus$p205p226_diet_ea <- rowSums(lupus[c("p205diet_ea", "p226diet_ea")])

# Ratio of dietary to total intake (after energy-adjustment)
lupus %>% 
  mutate(p183_diet_ratio     = p183diet_ea / p183_ea,
         n3pfa_diet_ratio    = n3pfa_diet_ea / n3pfa_ea,
         n6pfa_diet_ratio    = n6pfa_diet_ea / n6pfa_ea,
         p205p226_diet_ratio = p205p226_diet_ea / p205p226_ea) %>% 
  pivot_longer(ends_with("_ratio"), names_to = "var", values_to = "value") %>% 
  mutate(var = factor(var, levels = c("p183_diet_ratio", "p205p226_diet_ratio", "n3pfa_diet_ratio", "n6pfa_diet_ratio"))) %>% 
  group_by(var) %>% 
  summarize(mean = round(mean(value, na.rm = TRUE), 4)) %>% 
  knitr::kable()
```

* Intake of fatty acids (dietary and total) among those who use fish oil supplement:
  * Total intake of EPA + DHA is much greater, and most of its intake is from supplementary sources

```{r echo = FALSE}
# Fish oil supplement users only
table_vars <- c("p205p226_ea", "p205p226_diet_ea", "n3pfa_ea", "n3pfa_diet_ea")

lupus %>% 
  filter(take_fo == "Yes") %>% 
  set_column_labels(prev_sle          = "Prevalent SLE", 
                    p205p226_ea       = "DHA + EPA",
                    p205p226_diet_ea  = "DHA + EPA dietary",
                    n3pfa_ea          = "Omega-3",
                    n3pfa_diet_ea     = "Omega-3 dietary") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = prev_sle, 
                        header_count = "(n = %s)",
                        continuous_fn = describeMedian,
                        digits = 2,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) energy-adjusted intake of fatty acids (gram/day)",
            tfoot = "P-values were from Mann-Whitney tests")
```

* Three ratios of fatty acids, Omega-3/Omega-6, (DHA + EPA)/Omega-6, ALA/Omega-6, were calculated based on energy-adjusted intake, i.e., both the numerator and denominator were energy-adjusted values.
  * Distribution of the three ratios were highly right-skewed:
  
```{r echo = FALSE, fig.width = 10, fig.height = 3}
# Omega-3 and -6 fatty acids
table_vars <- c("o3_o6", "p205p226_o6", "p183_o6")

# Check distribution
lupus %>% 
  select(all_of(table_vars)) %>% 
  pivot_longer(o3_o6:p183_o6, names_to = "var", values_to = "value") %>% 
  mutate(var = factor(var, levels = table_vars))  %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~var, scales = "free", ncol = 3) +
  labs(x = "Ratio of energy-adjusted FA intake (gram/day)")
```

* A Spearman correlation matrix among three ratio variables is shown below.
  * Omega-3/omema-6 ratio is highly correlated with ALA/omega-6 ratio.

```{r echo = FALSE}
# Check Spearman correlations
lupus %>% 
  select(all_of(table_vars)) %>% 
  cor(method = "spearman") %>% 
  round(3) %>% 
  knitr::kable()
```

* Median (IQR) ratio of energy-adjusted fatty acids by cases/non-cases:
  * For all ratios, those with SLE had significantly higher intake than those without SLE.

```{r echo = FALSE}
table_vars <- c("o3_o6", "p205p226_o6", "p183_o6")
lupus %>% 
  set_column_labels(prev_sle    = "Prevalent SLE", 
                    o3_o6       = "Omega-3/Omega-6",
                    p205p226_o6 = "(DHA + EPA)/Omega-6",
                    p183_o6     = "ALA/Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = prev_sle, 
                        header_count = "(n = %s)",
                        continuous_fn = describeMedian,
                        digits = 4,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) ratio of energy-adjusted fatty acid intake",
            tfoot = "P-values were from Mann-Whitney tests")
```

## Quartiles of FA ratios

* The three FA ratio variables were categorized into quartile groups. Frequency tables by SLE status were shown below
  * For omega-3/omega-6 and (DHA + EPA)/omega-6, the SLE group had significantly more subjects in higher quartile groups, compared to the non-SLE group.

```{r echo = FALSE}
# Get quartiles of ratio variables
lupus <- lupus %>% 
  mutate(o3_o6       = n3pfa_ea / n6pfa_ea,
         p183_o6     = p183_ea / n6pfa_ea,
         p205p226_o6 = p205p226_ea / n6pfa_ea,
         o3_o6_cat       = ntile(o3_o6, 4),
         o3_o6_cat       = factor(o3_o6_cat),
         p183_o6_cat     = ntile(p183_o6, 4),
         p183_o6_cat     = factor(p183_o6_cat),
         p205p226_o6_cat = ntile(p205p226_o6, 4),
         p205p226_o6_cat = factor(p205p226_o6_cat))

# FA quartiles by SLE status
table_vars <- c("o3_o6_cat", "p205p226_o6_cat", "p183_o6_cat")
lupus %>% CreateTableOne(table_vars, strata = "prev_sle", data = .) %>%
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()
```

* Cut-off values of quartiles were shown below:

```{r echo = FALSE}
# Cut-off values
lupus %>% 
  select(starts_with("o3_o6")) %>% 
  group_by(o3_o6_cat) %>% 
  summarize(min = min(o3_o6), max = max(o3_o6)) %>% 
  knitr::kable()

lupus %>% 
  select(starts_with("p205p226_o6")) %>% 
  group_by(p205p226_o6_cat) %>% 
  summarize(min = min(p205p226_o6), max = max(p205p226_o6)) %>% 
  knitr::kable()

lupus %>% 
  select(starts_with("p183_o6")) %>% 
  group_by(p183_o6_cat) %>% 
  summarize(min = min(p183_o6), max = max(p183_o6)) %>% 
  knitr::kable()
```

## Logistic models with all three ratio variables

* Four logistic models were fitted using prevalent SLE as the outcome.
  * Model 1 includes: quartile groups of omega-3/omega-6, (DHA + EPA)/omega-6, ALA/omaga-6, and kcal (per 100 kcal)
  * Model 2 adds age groups, race, gender, education level, and smoking status (never/ever)
  * Model 3 further adds dietary pattern
  * Model 4 further adds BMI groups
* The table below shows estimated odds ratios and 95% Wald confidence intervals from each logistic model:

```{r echo = FALSE, results = 'asis'}
# Change references
lupus_md <- lupus %>% 
  mutate(kcal100  = kcal / 100,
         vegstat3 = relevel(vegstat3, ref = "Non-veg"),
         educat3  = relevel(educat3, ref = "Col grad"),
         agecat   = relevel(agecat, ref = ">=60"))

covar_labels <- c("Kcal/100",
                  "Age.: 30-39", 
                  "Age.: 40-59", 
                  "Race: Black",
                  "Sex.: Male",
                  "Educ: HS or less",
                  "Educ: Some college",
                  "Smkg: Ever", 
                  "Diet: Vegetarians",
                  "Diet: Pesco veg",
                  "BMI.: Overweight",
                  "BMI.: Obese")

# Function for stargazer
my_stargazer <- function(x){
  stargazer::stargazer(x, 
                       type = "html", 
                       digits = 2,
                       dep.var.caption = "",
                       dep.var.labels = "Outcome: Prevalent SLE",
                       model.numbers = FALSE,
                       column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
                       covariate.labels = var_labels,
                       apply.coef = exp, 
                       ci.custom = ci, 
                       star.cutoffs = NA, 
                       omit = "Constant", 
                       omit.stat = c("aic", "ll"),
                       omit.table.layout = "n")
}

fm <- formula(prev_sle ~ o3_o6_cat + p205p226_o6_cat + p183_o6_cat + I(kcal/100))
m1 <- glm(fm, family = binomial, data = lupus_md)
m2 <- update(m1, .~. + agecat + black + sex + educat3 + smkever)
m3 <- update(m2, .~. + vegstat3)
m4 <- update(m3, .~. + bmicat)

models <- list(m1, m2, m3, m4)
ci <- lapply(models, \(x) exp(confint.default(x)))

var_labels <- c("O3/O6: Q2",
                "O3/O6: Q3",
                "O3/O6: Q4",
                "DHA+EPA/O6: Q2",
                "DHA+EPA/O6: Q3",
                "DHA+EPA/O6: Q4",
                "ALA/O6: Q2",
                "ALA/O6: Q3",
                "ALA/O6: Q4",
                covar_labels)

my_stargazer(models)
```

## Logistic models with omega-3/omega-6 ratio

* Four logistic models were fitted using prevalent SLE as the outcome.
  * Model 1 includes: quartile groups of omega-3/omega-6 and kcal (per 100 kcal)
  * Model 2 adds age groups, race, gender, education level, and smoking status (never/ever)
  * Model 3 further adds dietary pattern
  * Model 4 further adds BMI groups
* The table below shows estimated odds ratios and 95% Wald confidence intervals from each logistic model:

```{r echo = FALSE, results = 'asis'}
# Logistic regression with omega-3/omega-6 ratio
fm <- formula(prev_sle ~ o3_o6_cat + I(kcal/100))
m1 <- glm(fm, family = binomial, data = lupus_md)
m2 <- update(m1, .~. + agecat + black + sex + educat3 + smkever)
m3 <- update(m2, .~. + vegstat3)
m4 <- update(m3, .~. + bmicat)

models <- list(m1, m2, m3, m4)
ci <- lapply(models, \(x) exp(confint.default(x)))

var_labels <- c("O3/O6: Q2",
                "O3/O6: Q3",
                "O3/O6: Q4",
                covar_labels)

my_stargazer(models)
```

* P-values for trend are shown below:

```{r echo = FALSE}
# Trend p-values
getLastPval <- function(x) broom::tidy(x) %>% select(p.value) %>% slice(nrow(.))

Mod1 <- Mod2 <- Mod3 <- Mod4 <- Mod5a <- Mod5b <- list()

# Model 1, ratio
Mod1["o3_o6_cat"] <- update(m1, .~. - o3_o6_cat + as.numeric(o3_o6_cat)) %>% 
  getLastPval()

# Model 2, ratio
Mod2["o3_o6_cat"] <- update(m2, .~. - o3_o6_cat + as.numeric(o3_o6_cat)) %>% 
  getLastPval()

# Model 2, age
Mod2["agecat"] <- update(m2, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 2, educ
Mod2["educat3"] <- update(m2, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, ratio
Mod3["o3_o6_cat"] <- update(m3, .~. - o3_o6_cat + as.numeric(o3_o6_cat)) %>% 
  getLastPval()

# Model 3, age
Mod3["agecat"] <- update(m3, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 3, educ
Mod3["educat3"] <- update(m3, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, vegstat3
Mod3["vegstat3"] <- update(m3, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

# Model 4, ratio
Mod4["o3_o6_cat"] <- update(m4, .~. - o3_o6_cat + as.numeric(o3_o6_cat)) %>% 
  getLastPval()

# Model 4, age
Mod4["agecat"] <- update(m4, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 4 educ
Mod4["educat3"] <- update(m4, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 4, vegstat3
Mod4["vegstat3"] <- update(m4, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

all_trend <- list(Mod1, Mod2, Mod3, Mod4)
names(all_trend) <- c("Model 1", "Model 2", "Model 3", "Model 4")

all_trend %>% 
  map(as.data.frame) %>% 
  map(\(x) round(x, 4))
```

## Logistic models with (DHA + EPA)/omega-6 ratio

* Four logistic models were fitted using prevalent SLE as the outcome.
  * Model 1 includes: quartile groups of (DHA + EPA)/omega-6 and kcal (per 100 kcal)
  * Model 2 adds age groups, race, gender, education level, and smoking status (never/ever)
  * Model 3 further adds dietary pattern
  * Model 4 further adds BMI groups
* The table below shows estimated odds ratios and 95% Wald confidence intervals from each logistic model:

```{r echo = FALSE, results = 'asis'}
# Logistic regression with (DHA + EPA)/omega-6 ratio
fm <- formula(prev_sle ~ p205p226_o6_cat + I(kcal/100))
m1 <- glm(fm, family = binomial, data = lupus_md)
m2 <- update(m1, .~. + agecat + black + sex + educat3 + smkever)
m3 <- update(m2, .~. + vegstat3)
m4 <- update(m3, .~. + bmicat)

models <- list(m1, m2, m3, m4)
ci <- lapply(models, \(x) exp(confint.default(x)))

var_labels <- c("DHA+EPA/O6: Q2",
                "DHA+EPA/O6: Q3",
                "DHA+EPA/O6: Q4",
                covar_labels)

my_stargazer(models)
```
* P-values for trend are shown below:

```{r echo = FALSE}
# Trend p-values
Mod1 <- Mod2 <- Mod3 <- Mod4 <- Mod5a <- Mod5b <- list()

# Model 1, ratio
Mod1["p205p226_o6_cat"] <- update(m1, .~. - p205p226_o6_cat + as.numeric(p205p226_o6_cat)) %>% 
  getLastPval()

# Model 2, ratio
Mod2["p205p226_o6_cat"] <- update(m2, .~. - p205p226_o6_cat + as.numeric(p205p226_o6_cat)) %>% 
  getLastPval()

# Model 2, age
Mod2["agecat"] <- update(m2, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 2, educ
Mod2["educat3"] <- update(m2, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, ratio
Mod3["p205p226_o6_cat"] <- update(m3, .~. - p205p226_o6_cat + as.numeric(p205p226_o6_cat)) %>% 
  getLastPval()

# Model 3, age
Mod3["agecat"] <- update(m3, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 3, educ
Mod3["educat3"] <- update(m3, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, vegstat3
Mod3["vegstat3"] <- update(m3, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

# Model 4, ratio
Mod4["p205p226_o6_cat"] <- update(m4, .~. - p205p226_o6_cat + as.numeric(p205p226_o6_cat)) %>% 
  getLastPval()

# Model 4, age
Mod4["agecat"] <- update(m4, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 4 educ
Mod4["educat3"] <- update(m4, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 4, vegstat3
Mod4["vegstat3"] <- update(m4, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

all_trend <- list(Mod1, Mod2, Mod3, Mod4)
names(all_trend) <- c("Model 1", "Model 2", "Model 3", "Model 4")

all_trend %>% 
  map(as.data.frame) %>% 
  map(\(x) round(x, 4))
```


## Logistic models with ALA/omega-6 ratio

* Four logistic models were fitted using prevalent SLE as the outcome.
  * Model 1 includes: quartile groups of ALA/omega-6 and kcal (per 100 kcal)
  * Model 2 adds age groups, race, gender, education level, and smoking status (never/ever)
  * Model 3 further adds dietary pattern
  * Model 4 further adds BMI groups
* The table below shows estimated odds ratios and 95% Wald confidence intervals from each logistic model:

```{r echo = FALSE, results = 'asis'}
# Logistic regression with ALA/omega-6 ratio
fm <- formula(prev_sle ~ p183_o6_cat + I(kcal/100))
m1 <- glm(fm, family = binomial, data = lupus_md)
m2 <- update(m1, .~. + agecat + black + sex + educat3 + smkever)
m3 <- update(m2, .~. + vegstat3)
m4 <- update(m3, .~. + bmicat)

models <- list(m1, m2, m3, m4)
ci <- lapply(models, \(x) exp(confint.default(x)))

var_labels <- c("ALA/O6: Q2",
                "ALA/O6: Q3",
                "ALA/O6: Q4",
                covar_labels)

my_stargazer(models)
```
* P-values for trend are shown below:

```{r echo = FALSE}
# Trend p-values
Mod1 <- Mod2 <- Mod3 <- Mod4 <- Mod5a <- Mod5b <- list()

# Model 1, ratio
Mod1["p183_o6_cat"] <- update(m1, .~. - p183_o6_cat + as.numeric(p183_o6_cat)) %>% 
  getLastPval()

# Model 2, ratio
Mod2["p183_o6_cat"] <- update(m2, .~. - p183_o6_cat + as.numeric(p183_o6_cat)) %>% 
  getLastPval()

# Model 2, age
Mod2["agecat"] <- update(m2, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 2, educ
Mod2["educat3"] <- update(m2, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, ratio
Mod3["p183_o6_cat"] <- update(m3, .~. - p183_o6_cat + as.numeric(p183_o6_cat)) %>% 
  getLastPval()

# Model 3, age
Mod3["agecat"] <- update(m3, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 3, educ
Mod3["educat3"] <- update(m3, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 3, vegstat3
Mod3["vegstat3"] <- update(m3, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

# Model 4, ratio
Mod4["p183_o6_cat"] <- update(m4, .~. - p183_o6_cat + as.numeric(p183_o6_cat)) %>% 
  getLastPval()

# Model 4, age
Mod4["agecat"] <- update(m4, .~. - agecat + as.numeric(agecat), data = lupus) %>% 
  getLastPval()

# Model 4 educ
Mod4["educat3"] <- update(m4, .~. - educat3 + as.numeric(educat3), data = lupus) %>% 
  getLastPval()

# Model 4, vegstat3
Mod4["vegstat3"] <- update(m4, .~. - vegstat3 + as.numeric(vegstat3), data = lupus) %>% 
  getLastPval()

all_trend <- list(Mod1, Mod2, Mod3, Mod4)
names(all_trend) <- c("Model 1", "Model 2", "Model 3", "Model 4")

all_trend %>% 
  map(as.data.frame) %>% 
  map(\(x) round(x, 4))
```

## Incident cases of SLE

```{r echo = FALSE}
lupus_inc <- lupus %>%
  filter(hhf3 == TRUE) %>% 
  filter(prev_sle == "No") %>% 
  mutate(inc_sle = if_else(!is.na(sle_dx), 1, 0),
         inc_sle = factor(inc_sle, labels = c("No", "Yes")))
```

* Incident cases of SLE were identified using [HHF3 Question 10](https://wiki.ahs2.org/_media/hhf3:hhf3-page-03.jpg).
  * There were *n* = `r lupus %>% filter(hhf3 == TRUE) %>% nrow()` subjects who returned HHF3 questionnaire. This included `r lupus %>% filter(hhf3 == TRUE, prev_sle == "Yes") %>% nrow()` prevalent cases at baseline.
  * The question asked when they were first diagnosed with SLE after 2001 in three time periods: 2002-2004, 2005-2006, and 2007-2008.
  * There were `r lupus %>% filter(!is.na(sle_dx)) %>% count()` participants who indicated their first diagnosis after 2001. However, after excluding prevalent cases, there were `r lupus %>% filter(!is.na(sle_dx), prev_sle == "No") %>% count()` incident cases of SLE. Among them, `r lupus %>% filter(!is.na(sle_dx), prev_sle == "No", sle_tx == "Yes") %>% count()` cases indicated that they were treated for SLE for the past 12 months.

* A frequency table of diagnosis periods:

```{r echo = FALSE}
lupus_inc %>% 
  group_by(sle_dx) %>% 
  filter(!is.na(sle_dx)) %>% 
  tally() %>% 
  mutate(Percent = round(n / sum(n) * 100, 1)) %>% 
  knitr::kable()
```

* Energy-adjusted FA intake was compared between cases and non-cases. 

```{r echo = FALSE}
# Compare FAs
table_vars <- c("p183_ea", "p205p226_ea", "n3pfa_ea", "n6pfa_ea")

# Table by SLE status
lupus_inc %>% 
  set_column_labels(inc_sle     = "Incident SLE", 
                    p183_ea     = "ALA",
                    p205p226_ea = "DHA + EPA",
                    n3pfa_ea    = "Omega-3",
                    n6pfa_ea    = "Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = inc_sle, 
                        continuous_fn = describeMedian,
                        header_count = "(n = %s)",
                        digits = 2,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) energy-adjusted intake of fatty acids (gram/day)",
            tfoot = "P-values were from Mann-Whitney tests")
```

* Ratios of energy-adjusted FA intake were compared between cases and non-cases. 

```{r echo = FALSE}
# Ratios
table_vars <- c("o3_o6", "p205p226_o6", "p183_o6")

# Table by SLE status
lupus_inc %>% 
  set_column_labels(inc_sle     = "Incident SLE", 
                    o3_o6       = "Omega-3/Omega-6",
                    p205p226_o6 = "(DHA + EPA)/Omega-6",
                    p183_o6     = "ALA/Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = inc_sle, 
                        continuous_fn = describeMedian,
                        header_count = "(n = %s)",
                        digits = 4,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) ratio of fatty acids",
            tfoot = "P-values were from Mann-Whitney tests")
```
