---
title: "AHS-2 lupus study 2"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "tableone", "emmeans", "kableExtra", "Gmisc")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, pattern, ...) {
  loc <- grep(pattern, names(df), ...)
  if (length(loc) == 0) warning("There are no such variables")
  else return(data.frame(loc = loc, varname = names(df)[loc]))
}

# Read data: n = 93,467 and 111 variables
filepath <- "./data/lupus-initial-dataset-v1-2022-04-25.csv"
lupus0  <- read_csv(filepath)
dim(lupus0)
names(lupus0)
n_distinct(lupus0$analysisid)

filepath <- "./data/lupus-initial-dataset-v2-2022-06-03.csv"
lupus00  <- read_csv(filepath)
dim(lupus00)

lupus0 <- lupus0 %>% 
  inner_join(select(lupus00, analysisid, starts_with("p182"), starts_with("p204")))

# Include non-Hispanic White or Black
# Exclude age < 30
# Exclude missing gender, diet pattern, education, smoking and BMI
# Exclude extreme kcal <500 or >4500
lupus <- lupus0 %>%
  mutate(
    age = ifelse(age < 30, NA, age),
    agecat = cut(age, breaks = c(0, 40, 60, Inf), right = FALSE),
    agecat = factor(agecat, labels = c("30-39", "40-59", ">=60")),
    sex = factor(sex, labels = c("Female", "Male")),
    black = case_when(
      ethyou == "01" ~ 0,
      ethyou %in% c("02", "03", "04", "05", "39", "40", "63", "74", "77", "96") ~ 1),
    black = factor(black, labels = c("White", "Black")),
    smkcat  = case_when(
      smokenow == 1 ~ 3,
      smoke > 1 & is.na(smokenow) ~ 2,
      smoke == 1 ~ 1
    ),
    smkcat  = factor(smkcat, labels = c("Never", "Past", "Current")),
    smkever = factor(smkcat, labels = c("Never", "Ever", "Ever")),
    educat3 = cut(educyou, breaks = c(0, 3, 6, 9)),
    educat3 = factor(educat3, labels = c("HS or less", "Some college", "Col grad")),
    # take_vd = ifelse(!is.na(vitd) & vitd == 2, 1, 0),
    # take_vd = factor(take_vd, labels = c("No", "Yes")),
    take_fo = ifelse(!is.na(fishoil) & fishoil == 2, 1, 0),
    take_fo = factor(take_fo, labels = c("No", "Yes")),
    bmicat  = cut(bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
    bmicat  = factor(bmicat, labels = c("Normal", "Overweight", "Obese")),
    vegstat = factor(vege_group_gen_bl, levels = c("vegan", "lacto", "pesco", "semi", "nonveg")),
    vegstat3 = vegstat,
    prev_sle = ifelse(!is.na(sle) & sle == 2, 1, 0),
    prev_sle = factor(prev_sle, labels = c("No", "Yes"))) %>% 
  drop_na(sex, age, black, vege_group_gen_bl, smkever, educat3, bmi) %>% 
  filter(kcal >= 500 & kcal <= 4500)

levels(lupus$vegstat) <- c("Vegans", "Lacto-ovo", "Pesco", "Semi", "Non-veg")
levels(lupus$vegstat3) <- c("Vegetarians", "Vegetarians", "Pesco", "Non-veg", "Non-veg")
```

## Dataset

* Filename: `lupus-initial-dataset-v1-2022-04-25.csv`
  * Data received on 4/25/2022
  * Includes *n* = `r nrow(lupus0)` subjects and `r ncol(lupus0)` variables
  * No imputations
  
* To create an analytic file:
  * Included only non-Hispanic White or Black
  * Excluded if age at baseline < 30 years
  * Excluded extreme energy intake of <500 or >4500 kcal
  * Excluded any subjects with missing gender, education, smoking history, dietary pattern and BMI
  * This resulted in *n* = `r nrow(lupus)` subjects (`r nrow(lupus0) - nrow(lupus)` subjects were excluded)

* **For now, both males and females are included in the analytic dataset**

## Outcome

* Prevalent cases of SLE were identified using the baseline questionnaire
  * Includes only those who "have been treated for SLE in the last 12 months" at baseline
  * ("Years since the 1st diagnosis" was not used, following the case definition in the manuscript)
* There were `r sum(lupus$prev_sle == "Yes")` prevalent cases (`r round(sum(lupus$prev_sle == "Yes") / nrow(lupus) * 100, 2)`%) of SLE
  
## Descriptive table

* Descriptive table stratified by cases/non-cases
  * Variables were categorized following the original manuscript

```{r echo = FALSE}
table_vars <- c("age", "agecat", "black", "sex", "smkever", "educat3", "vegstat3", "take_fo", "bmi", "bmicat")
lupus %>% CreateTableOne(table_vars, strata = "prev_sle", data = .) %>%
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()
```

## Energy-adjusted fatty acid intake

* The nutrient intake data include the following fatty acids
  * Each FA has both dietary and supplement intake

```{r echo = FALSE}
# Nutrients
fa <- tribble(
  ~fa_name, ~lipid, ~note,
  "ALA", "p183", "Omega-3",
  "SDA", "p184", "Omega-3",
  "EPA", "p205", "Omega-3",
  "DPA", "p225", "Omega-3",
  "DHA", "p226", "Omega-3",
  "LA",  "p182", "Omega-6",
  "AA",  "p204", "Omega-6"
)

fa %>% knitr::kable()
```

```{r echo = FALSE}
# Energy-adjust with zero partition
# By default, variables are log-transformed (excluding zeros)
kcal_adjust <- function(var, kcal, log=TRUE){
  if (missing(var))
    stop("Need to specify variable for energy-adjustment.")
  if (missing(kcal))
    stop("Need to specify kcal intake.")
  df <- data.frame(y = var, ea_y = var, kcal = kcal)
  count_negative <- sum(df$y < 0, na.rm=TRUE)
  if (count_negative > 0)
    warning("There are negative values in variable.")
  if(log) df$y[df$y > 0 & !is.na(df$y)] <- log(df$y[df$y > 0 & !is.na(df$y)])
  mod <- lm(y ~ kcal, data=df[df$y != 0, ])
  if(log){
    ea <- exp(resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE))
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  else{
    ea <- resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE)
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  return(df$ea_y)
}

# Create new variables
omega_vars         <- fa$lipid
omega_diet_vars    <- paste0(omega_vars, "diet")
omega_supp_vars    <- paste0(omega_vars, "supp")

omega_vars_ea      <- paste0(omega_vars, "_ea")
omega_diet_vars_ea <- paste0(omega_diet_vars, "_ea")

# Energy-adjust only dietary component (not supplement)
lupus[omega_diet_vars_ea] <- lapply(lupus[omega_diet_vars], kcal_adjust, kcal = lupus$kcal)

# Add energy-adjusted dietary intake and supplement
calc_total_ea <- function(var) rowSums(lupus[c(paste0(var, "diet_ea"), paste0(var, "supp"))])
lupus[omega_vars_ea] <- lapply(omega_vars, calc_total_ea)

# Sum omaga fa energy-adjusted values
lupus$n3pfa_ea    <- rowSums(lupus[c("p183_ea", "p184_ea", "p205_ea", "p225_ea", "p226_ea")])
lupus$n6pfa_ea    <- rowSums(lupus[c("p182_ea", "p204_ea")])
lupus$p205p226_ea <- rowSums(lupus[c("p205_ea", "p226_ea")])

# Calculate ratio of FA
lupus <- lupus %>% 
  mutate(o3_o6 = n3pfa_ea / n6pfa_ea,
         p183_o6 = p183_ea / n6pfa_ea,
         p205p226_o6 = p205p226_ea / n6pfa_ea)
```

* For each FA, its dietary intake was energy-adjusted by the residual method, while partitioning zero intake ([Jaceldo-Siegl et al., 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3433053/)). Due to its highly right-skewed distribution, non-zero intake was log-transformed before being regressed on the total energy intake. Residuals were added by the mean of log and then back-transformed to obtain energy-adjusted dietary intake. Energy-adjusted dietary intake was added with (unadjusted) supplement intake to form energy-adjuste total intake.
  * Energy-adjusted total omega-3 fatty acid intake was calculated as the sum of the following energy-adjusted FAs: ALA, SDA, EPA, DPA and DHA.
  * Energy-adjusted total omega-6 fatty acid intake was calculated as the sum of the following energy-adjusted FAs: LA, AA

* Median (IQR) intake (gram/day) of energy-adjusted ALA, EPA + DHA, total omega-3, and total omega-6 by cases/non-cases:
  * For EPA + DHA and total omega-3, those with SLE had significantly higher intake than those without SLE.

```{r echo = FALSE}
table_vars <- c("p183_ea", "p205p226_ea", "n3pfa_ea", "n6pfa_ea")
lupus %>% 
  set_column_labels(prev_sle    = "Prevalent SLE", 
                    p183_ea     = "ALA",
                    p205p226_ea = "DHA + EPA",
                    n3pfa_ea    = "Omega-3",
                    n6pfa_ea    = "Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = prev_sle, 
                        continuous_fn = describeMedian,
                        digits = 2,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) energy-adjusted intake of fatty acids (gram/day)",
            tfoot = "P-values were from Mann-Whitney tests")
```

* Median (IQR) ratio of energy-adjusted fatty acids by cases/non-cases:
  * For all ratios, those with SLE had significantly higher intake than those without SLE.

```{r echo = FALSE}
table_vars <- c("o3_o6", "p205p226_o6", "p183_o6")
lupus %>% 
  set_column_labels(prev_sle    = "Prevalent SLE", 
                    o3_o6       = "Omega-3/Omega-6",
                    p205p226_o6 = "(DHA + EPA)/Omega-6",
                    p183_o6     = "ALA/Omega-6") %>% 
  getDescriptionStatsBy(all_of(table_vars), 
                        by = prev_sle, 
                        continuous_fn = describeMedian,
                        digits = 4,
                        statistics = TRUE) %>% 
  htmlTable(caption = "Median (IQR) ratio of fatty acids",
            tfoot = "P-values were from Mann-Whitney tests")
```