---
title: "AHS-2 lupus study 2"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "tableone", "emmeans", "kableExtra")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, pattern, ...) {
  loc <- grep(pattern, names(df), ...)
  if (length(loc) == 0) warning("There are no such variables")
  else return(data.frame(loc = loc, varname = names(df)[loc]))
}

# Read data: n = 93,467 and 111 variables
filepath <- "./data/lupus-initial-dataset-v1-2022-04-25.csv"
lupus0  <- read_csv(filepath)

# Include non-Hispanic White or Black
# Exclude age < 30
# Exclude missing gender, diet pattern, education, smoking and BMI
# Exclude extreme kcal <500 or >4500
lupus <- lupus0 %>%
  mutate(
    age = ifelse(age < 30, NA, age),
    agecat = cut(age, breaks = c(0, 40, 60, Inf), right = FALSE),
    agecat = factor(agecat, labels = c("30-39", "40-59", ">=60")),
    sex = factor(sex, labels = c("Female", "Male")),
    black = case_when(
      ethyou == "01" ~ 0,
      ethyou %in% c("02", "03", "04", "05", "39", "40", "63", "74", "77", "96") ~ 1),
    black = factor(black, labels = c("White", "Black")),
    smkcat  = case_when(
      smokenow == 1 ~ 3,
      smoke > 1 & is.na(smokenow) ~ 2,
      smoke == 1 ~ 1
    ),
    smkcat  = factor(smkcat, labels = c("Never", "Past", "Current")),
    smkever = factor(smkcat, labels = c("Never", "Ever", "Ever")),
    educat3 = cut(educyou, breaks = c(0, 3, 6, 9)),
    educat3 = factor(educat3, labels = c("HS or less", "Some college", "Col grad")),
    # take_vd = ifelse(!is.na(vitd) & vitd == 2, 1, 0),
    # take_vd = factor(take_vd, labels = c("No", "Yes")),
    take_fo = ifelse(!is.na(fishoil) & fishoil == 2, 1, 0),
    take_fo = factor(take_fo, labels = c("No", "Yes")),
    bmicat  = cut(bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
    bmicat  = factor(bmicat, labels = c("Normal", "Overweight", "Obese")),
    vegstat = factor(vege_group_gen_bl, levels = c("vegan", "lacto", "pesco", "semi", "nonveg")),
    vegstat3 = vegstat,
    prev_sle = ifelse(!is.na(sle) & sle == 2, 1, 0),
    prev_sle = factor(prev_sle, labels = c("No", "Yes"))) %>% 
  drop_na(sex, age, black, vege_group_gen_bl, smkever, educat3, bmi) %>% 
  filter(kcal >= 500 & kcal <= 4500)

levels(lupus$vegstat) <- c("Vegans", "Lacto-ovo", "Pesco", "Semi", "Non-veg")
levels(lupus$vegstat3) <- c("Vegetarians", "Vegetarians", "Pesco", "Non-veg", "Non-veg")
```

## Dataset

* Filename: `lupus-initial-dataset-v1-2022-04-25.csv`
  * Data received on 4/25/2022
  * Includes *n* = `r nrow(lupus0)` subjects and `r ncol(lupus0)` variables
  * No imputations
  
* To create an analytic file:
  * Included only non-Hispanic White or Black
  * Excluded if age at baseline < 30 years
  * Excluded extreme energy intake of <500 or >4500 kcal
  * Excluded any subjects with missing gender, education, smoking history, dietary pattern and BMI
  * This resulted in *n* = `r nrow(lupus)` subjects (`r nrow(lupus0) - nrow(lupus)` subjects were excluded)

* **For now, both males and females are included in the analytic dataset**

## Outcome

* Prevalent cases of SLE were identified using the baseline questionnaire
  * Includes only those who "have been treated for SLE in the last 12 months" at baseline
  * ("Years since the 1st diagnosis" was not used, following the case definition in the manuscript)
* There were `r sum(lupus$prev_sle == "Yes")` prevalent cases (`r round(sum(lupus$prev_sle == "Yes") / nrow(lupus) * 100, 2)`%) of SLE
  
## Descriptive table

* Descriptive table stratified by cases/non-cases
  * Variables were categorized following the original manuscript

```{r echo = FALSE}
table_vars <- c("age", "agecat", "black", "sex", "smkever", "educat3", "vegstat3", "take_fo", "bmi", "bmicat")
lupus %>% CreateTableOne(table_vars, strata = "prev_sle", data = .) %>%
  print(showAllLevels = TRUE, pDigits = 4) %>% 
  kableone()
```

## Fatty acid intake

* Currently, the nutrient intake data include the following fatty acids
  * Each FA has both dietary and supplement intake
  * (What are omega-6 FA used in the manuscript? p182 + p204?)

```{r echo = FALSE}
# Nutrients
fa <- tribble(
  ~fa_name, ~lipid, ~note,
  "ALA", "p183", "Omega-3",
  "SDA", "p184", "Omega-3",
  "EPA", "p205", "Omega-3",
  "DPA", "p225", "Omega-3",
  "DHA", "p226", "Omega-3"
)

fa %>% knitr::kable()
```

* For each FA, its dietary intake was energy-adjusted by the residual method, while partitioning zero intake ([Jaceldo-Siegl et al., 2011](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3433053/)). Due to its highly right-skewed distribution, non-zero intake was log-transformed before being regressed on the total energy intake. Residuals were added by the mean of log and then back-transformed to obtain energy-adjusted dietary intake. Energy-adjusted dietary intake was added with (unadjusted) supplement intake to form energy-adjuste total intake.

* Energy-adjusted total omega-3 fatty acid intake was calculated as the sum of all energy-adjusted FAs (ALA, SDA, EPA, DPA and DHA).

* Median (IQR) intake of energy-adjusted ALA (`p183_ea`), EPA + DHA (`p205p226_ea`), and total omega-3 (`n3pfa_ea`) by cases/non-cases:
  * P-values were obtained with Mann-Whitney tests
  * For EPA + DHA and total omega-3, those with SLE had significantly higher intake than those without SLE.

```{r echo = FALSE}
# Energy-adjust with zero partition
# By default, variables are log-transformed (excluding zeros)
kcal_adjust <- function(var, kcal, log=TRUE){
  if (missing(var))
    stop("Need to specify variable for energy-adjustment.")
  if (missing(kcal))
    stop("Need to specify kcal intake.")
  df <- data.frame(y = var, ea_y = var, kcal = kcal)
  count_negative <- sum(df$y < 0, na.rm=TRUE)
  if (count_negative > 0)
    warning("There are negative values in variable.")
  if(log) df$y[df$y > 0 & !is.na(df$y)] <- log(df$y[df$y > 0 & !is.na(df$y)])
  mod <- lm(y ~ kcal, data=df[df$y != 0, ])
  if(log){
    ea <- exp(resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE))
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  else{
    ea <- resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE)
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  return(df$ea_y)
}

# Create new variables
o3_vars         <- fa$lipid
o3_diet_vars    <- paste0(o3_vars, "diet")
o3_supp_vars    <- paste0(o3_vars, "supp")

o3_vars_ea      <- paste0(o3_vars, "_ea")
o3_diet_vars_ea <- paste0(o3_diet_vars, "_ea")

# Energy-adjust only dietary component (not supplement)
lupus[o3_diet_vars_ea] <- lapply(lupus[o3_diet_vars], kcal_adjust, kcal = lupus$kcal)

# Add energy-adjusted dietary intake and supplement
calc_total_ea <- function(var) rowSums(lupus[c(paste0(var, "diet_ea"), paste0(var, "supp"))])
lupus[o3_vars_ea] <- lapply(o3_vars, calc_total_ea)

# Sum all omega-3 fatty acids energy-adjusted values
lupus$n3pfa_ea <- rowSums(lupus[o3_vars_ea])
lupus$p205p226_ea <- rowSums(lupus[c("p205_ea", "p226_ea")])
table_vars <- c("p183_ea", "p205p226_ea", "n3pfa_ea")
lupus %>% CreateTableOne(table_vars, strata = "prev_sle", data = .) %>%
  print(showAllLevels = TRUE, pDigits = 4, nonnormal = table_vars) %>% 
  kableone()
```

